Polytope notes

### 2020-01-24 notes/polytope-notes.tex
changes discussed with raf

  Useful lemmas:
  \begin{lemma} \label{lemma:D-polytope}
    For any $\T^r$, $D(\T^r)$ is a polytope (in particular, is convex).
  \end{lemma}
  \begin{proof}
    \bo{TODO}
  \end{proof}
  \begin{lemma}  \label{lemma:minkowski-support-all}
    The condition $\vec 0 \in \oplus_p \T^r$ holds if and only if, for all $w \in \reals^d$, $\sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w}{x} \geq 0$.
  \end{lemma}
  \begin{proof}
    Recall by definition, the notation $\oplus_p \T^r = \{\sum_y p(y) x^r_y : x^r_y \in T^r_y (\forall y)\}$.
    In particular, by the halfspace representation of this convex set, $x \in \oplus_p \T^r$ if and only if, for all $w$, $\inprod{w}{x} \leq \max_{x' \in \oplus_p \T^r} \inprod{w}{x'}$.
    Now, $\max_{x' \in \oplus_p \T^r} \inprod{w}{x'} = \sum_y p(y) \max_{x' \in T^r_y} \inprod{w}{x'}$.
    Plugging in $x = \vec{0}$ completes the proof.
  \end{proof}
  \begin{lemma} \label{lemma:minkowski-support-finite}
    Given $\T^r$, there exists a finite collection $w_1,\ldots,w_K \in \reals^d$ for some $K$ such that $\vec 0 \in \oplus_p \T^r$ if and only if, for all $i \in [K]$, $\sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \geq 0$.
  \end{lemma}
  \begin{proof}
    [Bo: TODO! The point: If we fix the support of $p$, there are only a finite number of normals that matter, regardless of which $p$ we pick. Varying over all finitely many choices of the support still gives a finite number of normals.]
  \end{proof}
  \begin{lemma} \label{lemma:E-to-B}
    Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
    Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
  \end{lemma}
  \begin{proof}
    [Bo: ? Should cite]
  \end{proof}

  \begin{theorem} \label{thm:vertex-halfspace-opt}
    Let the polytopes $\T^r = \{T^r_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
    We have $D(\T^r) = C$ if and only if both the halfspace and vertex conditions hold.
  \end{theorem}
  \begin{proof}
    ($\implies$)
    Suppose $D(\T^r) = C$.
    First, we note that the vertex condition is immediate: For all $j \in [\ell]$, $p^j \in C$ which gives $p^j \in D(\T^r)$.
    To show the halfspace condition is satisfied, we first construct a matrix $E$ such that $Ep \geq 0 \iff Bp \geq 0$, then use our construction to pick out the necessary vectors $v_1,\dots,v_k$.

    By Lemma \ref{lemma:minkowski-support-finite}, there is a finite collection of vectors $w_1,\dots,w_{K} \in \reals^d$ where $\vec 0 \in \oplus_p \T^r$ if and only if, for all $w_i$, $\sum_y p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \geq 0$.
    Hence, each vector $w_i$ generates a row of a matrix $E \in \reals^{K \times n}$ with $E_{iy} = \max_{x \in T^r_y} \inprod{w_i}{x}$, and we have $0 \in \oplus_p \T^r \iff Ep \geq 0$.
    By assumption of $\Opt(\T^r,C)$, then, we have $Ep \geq 0 \iff Bp \geq 0$.
    By Lemma \ref{lemma:E-to-B}, because $B$ has the minimum possible number of rows, each row of $B$ appears (scaled by some positive constant) as a different row of $E$. Taking the collection of $w_i$ corresponding to these rows and rescaling them by that positive constant, we get a collection of $k$ vectors $v_1,\ldots,v_k \in \reals^d$ with $\max_{x \in T^r_y} \inprod{v_i}{x} = B_{iy}$, hence the halfspace condition is satisfied.

>>>>

  Useful lemmas:
  \begin{lemma} \label{lemma:D-polytope}
    For any $\T^r$, $D(\T^r)$ is a polytope (in particular, is convex).
  \end{lemma}
  \begin{proof}
%    Recall by definition, the notation $\oplus_p \T^r = \{\sum_y p(y) x^r_y : x^r_y \in T^r_y (\forall y)\}$.
%    Each $T^r_y$ is a polytope, so $p(y) T^r_y$ is a polytope.
%    The Minkowski sum of polytopes is a polytope, so $\oplus_p \T^r$ is a polytope.
    \bo{TODO}
  \end{proof}
  \begin{lemma}  \label{lemma:minkowski-support}
    Given polytopes $\T^r$, there exists a finite set of normal vectors $w_1,\ldots,w_K \in \reals^d$ such that, for all $p \in \simplex$, $\oplus_p \T^r = \{x : \inprod{w_i}{x} \leq \sum_{y \in \Y} p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \}$.
  \end{lemma}
  \begin{proof}
%    For each $p$, $\oplus_p \T^r$ is a polytope.
%
    \bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
  \end{proof}
  \begin{lemma} \label{lemma:E-to-B}
    Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
    Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
  \end{lemma}
  \begin{proof}
    \bo{? Cite}
  \end{proof}

  \begin{theorem} \label{thm:vertex-halfspace-opt}
    Let the polytopes $\T^r = \{T^r_y \subseteq \reals^d : y \in \Y\}$ and $C$ be given, with $C = \conv(\{p^1,\ldots,p^{\ell}\}) = \{p: Bp \geq \vec 0\}$ for $B \in \reals^{k \times n}$.
    We have $D(\T^r) = C$ if and only if both the halfspace and vertex conditions hold.
  \end{theorem}
  \begin{proof}
    ($\implies$)
    Suppose $D(\T^r) = C$.
    First, we note that the vertex condition is immediate: For all $j \in [\ell]$, $p^j \in C$ which gives $p^j \in D(\T^r)$.
    To show the halfspace condition is satisfied, we first construct a matrix $E$ such that $Ep \geq 0 \iff Bp \geq 0$, then use this construction to pick out the necessary vectors $v_1,\dots,v_k$.

    By Lemma \ref{lemma:minkowski-support}, there is a finite collection of vectors $w_1,\dots,w_{K} \in \reals^d$ and such that $\vec 0 \in \oplus_p \T^r$ if and only if, for all $w_i$, $\sum_y p(y) \max_{x \in T^r_y} \inprod{w_i}{x} \geq 0$.
    Hence, each vector $w_i$ generates a row of a matrix $E \in \reals^{K \times n}$ with $E_{iy} = \max_{x \in T^r_y} \inprod{w_i}{x}$, and we have $p \in D(\T^r) \iff Ep \geq 0$.
    By assumption of $D(\T^r) = C$, then, we have $Ep \geq 0 \iff Bp \geq 0$.
    By Lemma \ref{lemma:E-to-B}, because $B$ has the minimum possible number of rows, each row of $B$ appears (scaled by some positive constant) as a different row of $E$. Taking the collection of $w_i$ corresponding to these rows and rescaling them by that positive constant, we get a collection of $k$ vectors that we can rename $v_1,\ldots,v_k \in \reals^d$, with $\max_{x \in T^r_y} \inprod{v_i}{x} = B_{iy}$, hence the halfspace condition is satisfied.
